---
title: "Initial Processing"
author: "Dariia Artemova"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    df_print: "paged"
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(Seurat)
library(Azimuth)
here::i_am("code/01_initial_processing.Rmd")
library(here)
knitr::opts_chunk$set(
  fig.align = 'center'
)
```

## Loading Data

```{r load-counts}
counts <- read.delim(
  file = here("data/raw/GBM_raw_gene_counts.csv"), 
  sep = " ", 
  row.names = 1, 
  check.names = F,
  ) 
head(counts)[c(1:15, ncol(counts)-5:ncol(counts))]
```

We loaded `r ncol(counts)` cells and `r nrow(counts)` genes.

```{r load-metadata}
metadata <- read.delim(
  file = here("data/raw/GBM_metadata.csv"), 
  sep = " ", 
  row.names = 1, 
  stringsAsFactors = TRUE,
)
head(metadata)
```

```{r tidy-metadata}
metadata <- metadata |> 
  select(!c(Sample.type, housekeeping_cluster, ends_with("color"))) |> 
  rename(Patient = Sample.name, Paper_cluster = Cluster_2d) |> 
  unite("Batch", Patient, Location, remove = FALSE) |> 
  relocate(Batch, Patient, Location, Selection) |> 
  mutate(Batch = as_factor(Batch))
head(metadata, n = 3)
```

```{r}
table(metadata[c("Patient", "Location")])
```

```{r patient-data}
patient_data <- read_excel(
  here("data/raw/patient_data.xlsx"),
  range = "A2:V6",
  .name_repair = "unique_quiet"
  ) |> 
  rename(Patient = 1) |>
  rename_with(~ gsub(" ", "_", .x, fixed = TRUE)) |>
  rename_with(~ paste("Subtype", .x, sep = "_"), Classical:Proneural) |>
  rename_with(~ paste(.x, "expressing_cell_fraction", sep = "_"), CD274:B2M)
patient_data
```

Where WT = Wildtype, M = Methylated, NM = Non methylated, NT = Not tested.

```{r seurat-obj, warning=FALSE}
options(Seurat.object.assay.calcn = TRUE)
obj <- CreateSeuratObject(
  counts = counts,
  meta.data = metadata,
  min.cells = 3,
  min.features = 100,
  )
```

We're left with `r length(Cells(obj))` cells and `r length(Features(obj))` genes.

## QC

```{r feature-counts-vis, warning=FALSE}
VlnPlot(
  obj,
  features = c("nFeature_RNA", "nCount_RNA", "ERCC_reads"),
  layer = "counts",
) &
  labs(x = NULL) &
  scale_x_discrete(labels = NULL, breaks = NULL)
```

```{r spike-in-vis, warning=FALSE, fig.width=3}
VlnPlot(
  obj,
  features = "ERCC_to_non_ERCC",
  layer = "counts",
  y.max = 3,
) +
  labs(x = NULL) +
  scale_x_discrete(labels = NULL, breaks = NULL) +
  guides(fill="none")
```

```{r qc-feature-scatter}
n_feature_cutoff = 200
n_count_cutoff = 50000

FeatureScatter(
  obj,
  feature1 = "nCount_RNA", 
  feature2 = "nFeature_RNA",
  cols = "black"
) +
  NoLegend() +
  geom_hline(
    aes(yintercept = n_feature_cutoff), 
    color = "red", linetype = "dashed", 
  ) +
  geom_vline(
    aes(xintercept = n_count_cutoff), 
    color = "red", linetype = "dashed",
  )
```

```{r, include=FALSE}
n_cells_pre <- length(Cells(obj))
```

```{r qc-subsetting}
obj <- subset(obj, subset = nFeature_RNA > n_feature_cutoff & nCount_RNA > n_count_cutoff)
```

`r n_cells_pre - length(Cells(obj))` cells filtered out during QC.

## Cell Type Annotation

```{r azimuth-annotation, include=FALSE}
options(future.globals.maxSize = 1000 * 1024^2)
obj <- RunAzimuth(obj, reference = "humancortexref", verbose = F)
```

```{r gen-vis, fig.width=15, fig.height=10}
plot_cell_cls <- function(reduction) {
  (DimPlot(
  obj, 
  reduction = reduction,
  group.by = "predicted.subclass",
  cols = DiscretePalette(n = length(unique(obj$predicted.subclass)), palette = "polychrome"),
  pt.size = 0.5,
) | 
  FeaturePlot(
    obj, 
    reduction = reduction,
    features = "predicted.subclass.score",
    pt.size = 0.5,
  ))  / 
  (DimPlot(
    obj, 
    reduction = reduction,
    group.by = "Batch",
    cols = DiscretePalette(n = nlevels(obj$Batch), palette = "polychrome"),
    pt.size = 0.5,
  ) |
  DimPlot(
      obj, 
      reduction = reduction,
      group.by = "Selection",
      cols = DiscretePalette(n = nlevels(obj$Selection), palette = "polychrome"),
      pt.size = 0.5,
  ))
}
plot_cell_cls("ref.umap")
```

## Transformation / DimRed (No Integration Version)

```{r no-intgr-pca}
obj <- SCTransform(object = obj, verbose = FALSE) |> 
  RunPCA(verbose = FALSE)
ElbowPlot(object = obj, ndims = 50)
```

```{r no-intgr-umap, warning=FALSE, fig.width=15}
obj <- RunUMAP(object = obj, dims = 1:30, verbose = FALSE)

DimPlot(
  object = obj, 
  reduction = "umap",
  group.by = "Batch",
  cols = DiscretePalette(n = nlevels(obj$Batch), palette = "polychrome"),
  pt.size = 0.1,
) + 
  DimPlot(
    object = obj, 
    reduction = "umap",
    group.by = "Selection",
    pt.size = 0.1,
  )
```

## Transformation / DimRed (With Integration This Time)

```{r integration, message=FALSE, warning=FALSE}
obj[["RNA"]] <- split(obj[["RNA"]], f = obj$Batch)
obj <- SCTransform(object = obj, verbose = FALSE)
obj <- RunPCA(object = obj, verbose = FALSE)
ElbowPlot(object = obj, ndims = 50)
obj <- IntegrateLayers(
  object = obj,
  method = HarmonyIntegration,
  normalization.method = "SCT",
  orig.reduction = "pca", new.reduction = "harmony",
  assay = "SCT",
  verbose = FALSE
)
```

```{r integration-umap, message=FALSE, warning=FALSE}
obj <- RunUMAP(object = obj, reduction = "harmony", dims = 1:30, reduction.name = "umap_harmony", verbose = F)
```

```{r harmony-vis, fig.width=15, fig.height=10}
plot_cell_cls("umap_harmony")
```

## Saving Results

```{r save-rds}
saveRDS(
   object = obj,
   file = here("data/processed/01_initial_processing.Rds")
)
```

<details>
  <summary>Session Info</summary>
```{r session-info}
sessionInfo()
```
</details>

