---
title: "Initial Processing"
output: 
  html_document:
    df_print: "paged"
date: "2025-04-02"
---

```{r libraries, include=FALSE}
library(tidyverse)
library(readxl)
library(Seurat)
library(dplyr)
```

## TODO:

-   properly validate this nonsense
-   better QC
-   annotate cell types
-   find doublets
-   PaCMAP

## Loading Data

```{r load-counts}
counts <- read.delim(
  file = "GBM_raw_gene_counts.csv", 
  sep = " ", 
  row.names = 1, 
  check.names = F,
  ) 
head(counts)[c(1:15, ncol(counts)-5:ncol(counts))]
```

We loaded `r ncol(counts)` cells and `r nrow(counts)` genes.

```{r load-metadata}
metadata <- read.delim(
  file = "GBM_metadata.csv", 
  sep = " ", 
  row.names = 1, 
  stringsAsFactors = TRUE,
)
head(metadata)
```

```{r tidy-metadata}
metadata <- metadata |>
  select(!c(Sample.type, housekeeping_cluster, ends_with("color"))) |>
  rename(Patient = Sample.name, Paper_cluster = Cluster_2d) |>
  unite("Batch", Patient, Location, remove = FALSE) |>
  relocate(Batch, Patient, Location, Selection) |>
  mutate(Batch = as_factor(Batch))
head(metadata, n = 3)
```

```{r}
table(metadata[c("Patient", "Location")])
```

```{r patient-data}
patient_data <- read_excel(
  "patient_data.xlsx",
  range = "A2:V6",
  .name_repair = "unique_quiet"
  ) |> 
  rename(Patient = 1) |>
  rename_with(~ gsub(" ", "_", .x, fixed = TRUE)) |>
  rename_with(~ paste("Subtype", .x, sep = "_"), Classical:Proneural) |>
  rename_with(~ paste(.x, "expressing_cell_fraction", sep = "_"), CD274:B2M)
patient_data
```

Where WT = Wildtype, M = Methylated, NM = Non methylated, NT = Not tested.

```{r seurat-obj, warning=FALSE}
obj <- CreateSeuratObject(
  counts = counts,
  meta.data = metadata,
  min.cells = 3,
  min.features = 100,
  )
```

We're left with `r length(Cells(obj))` cells and `r length(Features(obj))` genes.

## QC

```{r feature-counts-vis, warning=FALSE, fig.align='center'}
VlnPlot(
  obj,
  features = c("nFeature_RNA", "nCount_RNA", "ERCC_reads"),
  layer = "counts",
) &
  labs(x = NULL) &
  scale_x_discrete(labels = NULL, breaks = NULL)
```

```{r spike-in-vis, warning=FALSE, fig.width=3, fig.align='center'}
VlnPlot(
  obj,
  features = "ERCC_to_non_ERCC",
  layer = "counts",
  y.max = 3,
) +
  labs(x = NULL) +
  scale_x_discrete(labels = NULL, breaks = NULL) +
  guides(fill="none")
```

```{r qc-feature-scatter, fig.align='center'}
n_feature_cutoff = 200
n_count_cutoff = 50000

FeatureScatter(
  obj,
  feature1 = "nCount_RNA", 
  feature2 = "nFeature_RNA",
  cols = "black"
) +
  NoLegend() +
  geom_hline(
    aes(yintercept = n_feature_cutoff), 
    color = "red", linetype = "dashed", 
  ) +
  geom_vline(
    aes(xintercept = n_count_cutoff), 
    color = "red", linetype = "dashed",
  )
```

```{r, include=FALSE}
n_cells_pre <- length(Cells(obj))
```

```{r qc-subsetting}
obj <- subset(obj, subset = nFeature_RNA > n_feature_cutoff & nCount_RNA > n_count_cutoff)
```

`r n_cells_pre - length(Cells(obj))` cells filtered out during QC.

## Transformation / DimRed (No Integration Version)

```{r no-intgr-pca}
obj <- SCTransform(object = obj, verbose = FALSE) |> 
  RunPCA(verbose = FALSE)
ElbowPlot(object = obj, ndims = 50)
```

```{r no-intgr-umap, warning=FALSE, fig.width=15, fig.align='center'}
obj <- RunUMAP(object = obj, dims = 1:30, verbose = FALSE)

DimPlot(
  object = obj, 
  reduction = "umap",
  group.by = "Batch",
  cols = DiscretePalette(n = nlevels(obj$Batch), palette = "polychrome"),
  pt.size = 0.1,
) + 
  DimPlot(
    object = obj, 
    reduction = "umap",
    group.by = "Selection",
    pt.size = 0.1,
  )
```

## Transformation / DimRed (With Integration This Time)

```{r integration, message=FALSE, warning=FALSE, fig.align='center'}
obj[["RNA"]] <- split(obj[["RNA"]], f = obj$Batch)
obj <- SCTransform(object = obj, verbose = FALSE)
obj <- RunPCA(object = obj, verbose = FALSE)
ElbowPlot(object = obj, ndims = 50)
obj <- IntegrateLayers(
  object = obj,
  method = HarmonyIntegration,
  normalization.method = "SCT",
  orig.reduction = "pca", new.reduction = "harmony",
  assay = "SCT",
  verbose = FALSE
)
```

```{r integration-umap, message=FALSE, warning=FALSE, fig.width=15, fig.align='center'}
obj <- RunUMAP(object = obj, reduction = "harmony", dims = 1:30, reduction.name = "umap_harmony", verbose = F)

DimPlot(
  object = obj, 
  reduction = "umap_harmony",
  group.by = "Batch",
  cols = DiscretePalette(n = nlevels(obj$Batch), palette = "polychrome"),
  pt.size = 0.1,
) + 
  DimPlot(
    object = obj, 
    reduction = "umap_harmony",
    group.by = "Selection",
    pt.size = 0.1,
  )
```

```{r}
obj <- SCTransform(object = obj, verbose = FALSE)
obj <- RunPCA(object = obj, verbose = FALSE)
obj <- FindNeighbors(obj, dims = 1:30)
obj <- FindClusters(obj, resolution = 0.6)
obj <- RunUMAP(object = obj, reduction = "harmony", dims = 1:30, reduction.name = "umap_harmony", verbose = F)

DimPlot(obj, reduction = "umap_harmony", group.by = "Selection", label = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()
DimPlot(obj, reduction = "umap_harmony", group.by = "Location", label = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()
DimPlot(obj, reduction = "umap_harmony", label = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()
```

```{r}
FeaturePlot(obj, features = "PTPRC", reduction = "umap_harmony", pt.size = 0.1)
FeaturePlot(obj, features = "OLIG1", reduction = "umap_harmony", pt.size = 0.1)
```

```{r}
obj <- PrepSCTFindMarkers(obj)
markers <- FindMarkers(obj, ident.1 = "0", ident.2 = "2", min.diff.pct = 0.25,test.use = "wilcox")
```

```{r}
markers %>% filter(abs(avg_log2FC) > 0.5, p_val_adj < 0.05) %>% arrange(desc(avg_log2FC))
```

```{r warning=FALSE}
cns_markers <- list(
  Astrocytes = c("GFAP", "AQP4", "ALDH1L1", "S100B", "SLC1A3"),
  Astrocytes_Protoplasmic = c("SLC1A2", "ALDH1L1", "GFAP"),
  Astrocytes_Fibrous = c("GFAP", "VIM", "CD44"),
  Oligodendrocyte_Precursors = c("PDGFRA", "CSPG4"),
  Oligodendrocytes_Mature = c("MBP", "MOG", "PLP1"),
  Neurons_Excitatory = c("SLC17A7", "CAMK2A", "GRIN1"),
  Neurons_Inhibitory = c("GAD1", "GAD2", "SLC32A1"),
  Microglia = c("TMEM119", "P2RY12", "HEXB", "SALL1", "GPR34", "OLFML3", "MERTK", "PROS1", "TYRO3", "TGFBR1"))

obj <- AddModuleScore(object = obj,
  features = cns_markers,
  name = "CNS_Score"
)

score_columns <- grep("CNS_Score", colnames(obj@meta.data), value = TRUE)
obj@meta.data$assigned_celltype <- factor(apply(obj@meta.data[, score_columns], 1, function(x) names(cns_markers)[which.max(x)]))
```

```{r}
cluster_markers <- FindAllMarkers(
    obj, only.pos = TRUE,
    min.pct = 0.25, logfc.threshold = 0.25,
    group.by = "seurat_clusters"
)

celltype_markers <- FindAllMarkers(
    obj, only.pos = TRUE,
    min.pct = 0.25, logfc.threshold = 0.25,
    group.by = "assigned_celltype"
)
```

```{r}
celltype_markers
```

```{r}
celltype_markers %>% mutate(pct_diff = pct.1 - pct.2) %>% filter(p_val_adj < 0.05, avg_log2FC > 1, pct_diff > 0.3) %>% filter(cluster == "Oligodendrocyte_Precursors")
```

```{r}
celltype_markers %>% mutate(pct_diff = pct.1 - pct.2) %>% filter(p_val_adj < 0.05, avg_log2FC > 1, pct_diff > 0.3) %>% filter(cluster == "Oligodendrocytes_Mature")
```

```{r}
celltype_markers %>% mutate(pct_diff = pct.1 - pct.2) %>% filter(p_val_adj < 0.05, avg_log2FC > 1, pct_diff > 0.3) %>% filter(cluster == "Microglia")
```

```{r}
top5_clusters <- cluster_markers %>% group_by(cluster) %>% filter(p_val_adj < 0.05, avg_log2FC > 1) %>% top_n(n = 3, wt = avg_log2FC) 
top5_celltypes <- celltype_markers %>% group_by(cluster) %>% filter(p_val_adj < 0.05, avg_log2FC > 1)  %>% top_n(n = 5, wt = avg_log2FC)

DoHeatmap(subset(obj, downsample = 30), features = top5_clusters$gene, assay = "RNA", slot = "counts") + NoLegend()
DoHeatmap(subset(obj, downsample = 30), features = top5_celltypes$gene, group.by = "assigned_celltype", assay = "RNA", slot = "counts") + NoLegend()
```

```{r}
obj@meta.data <- obj@meta.data %>% mutate(condition_celltype = paste0(Location, "_", assigned_celltype))
```

```{r}
unique(obj@meta.data$condition_celltype)
```

```{r}
Idents(obj) <- obj@meta.data$condition_celltype
```

```{r}
microglia_obj <- subset(obj, subset = assigned_celltype == "Microglia")

microglia_obj@meta.data$Tumor_Periphery <- ifelse(
  microglia_obj@meta.data$Location == "Tumor",
  "Tumor",
  ifelse(
    microglia_obj@meta.data$Location == "Periphery",
    "Periphery",
    ifelse(
      microglia_obj@meta.data$Location == "Distant",
      "Distant",
      "Other"
    )
  )
)
```

```{r warning=FALSE}
microglia_obj <- SCTransform(microglia_obj, verbose = FALSE)
microglia_obj <- PrepSCTFindMarkers(microglia_obj)
```

```{r}
# Compare Tumor vs. Periphery
Idents(microglia_obj) <- "Tumor_Periphery"
tumor_vs_periphery_genes <- FindMarkers(microglia_obj, ident.1 = "Tumor", ident.2 = "Periphery", min.pct = 0.25, logfc.threshold = 0.25)
```

```{r}
# Then, compare Tumor vs. Distant
tumor_vs_distant_genes <- FindMarkers(microglia_obj, ident.1 = "Tumor", ident.2 = "Distant", min.pct = 0.25, logfc.threshold = 0.25)
```

```{r}
# And, compare Periphery vs. Distant
periphery_vs_distant_genes <- FindMarkers(microglia_obj, ident.1 = "Periphery", ident.2 = "Distant", min.pct = 0.25, logfc.threshold = 0.25)
```

```{r}
#if (!requireNamespace("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")
#BiocManager::install("EnhancedVolcano")
library(EnhancedVolcano)

#if (!requireNamespace("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")
#BiocManager::install("clusterProfiler")
#BiocManager::install("org.Hs.eg.db")
library(clusterProfiler)
library(org.Hs.eg.db)
#if (!requireNamespace("BiocManager", quietly = TRUE))
  #install.packages("BiocManager")
#BiocManager::install("msigdbr")
library(msigdbr)
#if (!requireNamespace("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")
#BiocManager::install("enrichplot")
library(enrichplot)
```

```{r}
# Visualization and Interpretation

# Function to visualize and interpret DE results
visualize_de_results <- function(de_genes, comparison_name, top_n_heatmap = 20, top_n_features = 3) {
  print(paste("Results for:", comparison_name))

  # View top differentially expressed genes
  head(de_genes)

  # Filter for significant and substantial changes
  significant_genes <- de_genes %>%
    filter(p_val_adj < 0.05, abs(avg_log2FC) > 0.5) %>%
    arrange(desc(avg_log2FC))

  print(significant_genes)

  # Volcano Plot
  volcano_plot <- EnhancedVolcano::EnhancedVolcano(de_genes,
    lab = rownames(de_genes),
    x = "avg_log2FC",
    y = "p_val_adj",
    pCutoff = 0.05,
    FCcutoff = 0.5,
    title = paste("Volcano Plot:", comparison_name),
    subtitle = "Значущі гени виділено",
    caption = paste0("Всього змінних = ", nrow(de_genes)),
    col = c("grey30", "forestgreen", "royalblue", "firebrick2"),
    colAlpha = 0.8,
    boxedLabels = TRUE,
    drawConnectors = TRUE,
    widthConnectors = 0.5,
    colConnectors = "grey50",
    max.overlaps = 20,
    legendPosition = "topright",
    legendLabSize = 10,
    legendIconSize = 4.0,
    titleLabSize = 14,
    subtitleLabSize = 12,
    captionLabSize = 10,
    xlab = bquote(~Log[2]~ " зміна експресії"),
    ylab = bquote(~-Log[10]~italic(P)),
    pointSize = 1.5
  )
  print(volcano_plot)


  # Feature Plots (Top N genes)
  if (nrow(significant_genes) > 0) {
    tryCatch({
      top_feature_genes <- head(rownames(significant_genes), top_n_features)
      feature_plot <- FeaturePlot(microglia_obj, features = top_feature_genes, reduction = "umap_harmony", cols = c("lightgrey", "red"), pt.size = 0.5)
      print(feature_plot)
    }, error = function(e) {
      print(paste("Помилка при створенні FeaturePlot:", e$message))
    })
  }

# Heatmap Visualization (з групуванням за поточними Idents)
if (nrow(significant_genes) > 0) {
  # Вибираємо топ N значущих генів
  top_genes <- head(rownames(significant_genes), top_n_heatmap)

  # Масштабуємо дані за допомогою ScaleData
  microglia_obj <- ScaleData(microglia_obj, features = top_genes, verbose = FALSE)

  # Створюємо хітмапу (групування за поточними Idents)
  heatmap_plot <- DoHeatmap(object = microglia_obj, features = top_genes, assay = "SCT", slot = "scale.data") +
    ggplot2::ggtitle(paste("Heatmap (Scaled):", comparison_name)) +
    ggplot2::theme(plot.title = ggplot2::element_text(color = "white", size = 2, hjust = 0, vjust = 0)) 
  print(heatmap_plot)
}


  gene_list <- rownames(significant_genes)
  if (length(gene_list) > 0) {
    entrez_ids <- bitr(gene_list, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Hs.eg.db)

    if (nrow(entrez_ids) > 0) {
      # GO Enrichment
      go_results <- enrichGO(gene = entrez_ids$ENTREZID,
        OrgDb = org.Hs.eg.db,
        ont = "BP",
        pAdjustMethod = "BH",
        pvalueCutoff = 0.01,
        qvalueCutoff = 0.05
      )

    
      
      if (!is.null(go_results)) {
  tryCatch({
    print(enrichplot::dotplot(go_results, showCategory = 10, title = paste("GO Enrichment:", comparison_name)) +
            ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8, angle = 0, hjust = 1, vjust = 0.5),
                           plot.title = ggplot2::element_text(hjust = 0.5))) 
  }, error = function(e) {
    print(paste("Недостатньо даних для створення dotplot або структура go_results не відповідає очікуваній.", e$message))
  })
}

      # Hallmark Gene Sets
library(msigdbr)
hallmark_gene_sets <- msigdbr(species = "Homo sapiens", category = "H")

gene_ranks_symbol <- de_genes$avg_log2FC
names(gene_ranks_symbol) <- rownames(de_genes)
entrez_map <- bitr(names(gene_ranks_symbol), fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Hs.eg.db)

# Створюємо ранжований список з Entrez ID як іменами
gene_ranks_entrez_named <- gene_ranks_symbol[entrez_map$SYMBOL]
names(gene_ranks_entrez_named) <- entrez_map$ENTREZID

# Видаляємо записи з NA в іменах (Entrez ID)
gene_ranks_entrez_filtered <- gene_ranks_entrez_named[!is.na(names(gene_ranks_entrez_named))]

# Сортуємо ранжований список
gene_ranks_entrez_filtered <- sort(gene_ranks_entrez_filtered, decreasing = TRUE)

hallmark_results <- GSEA(gene_ranks_entrez_filtered,
                         TERM2GENE = hallmark_gene_sets %>% dplyr::select(gs_name, entrez_gene),
                         pvalueCutoff = 0.05
)
if (!is.null(hallmark_results) && nrow(hallmark_results) > 0) {
  tryCatch({
    print(enrichplot::dotplot(hallmark_results, showCategory = 10, title = paste("Hallmark Enrichment:", comparison_name)) +
            ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8, angle = 0, hjust = 1, vjust = 0.5), 
                           plot.title = ggplot2::element_text(hjust = 0.5))) 
  }, error = function(e) {
    print(paste("Помилка при створенні dotplot (Hallmark):", e$message))
  })
} else {
  print("Немає значущих результатів збагачення Hallmark gene sets для відображення.")
}
    }
  }

}
```

```{r warning=FALSE}
# Apply visualization function to each comparison
visualize_de_results(tumor_vs_periphery_genes, "Tumor vs. Periphery (Microglia)", top_n_heatmap = 15, top_n_features = 4)
visualize_de_results(tumor_vs_distant_genes, "Tumor vs. Distant (Microglia)", top_n_heatmap = 15, top_n_features = 4)
visualize_de_results(periphery_vs_distant_genes, "Periphery vs. Distant (Microglia)", top_n_heatmap = 15, top_n_features = 4)
```

```{r}
oligo_obj <- subset(obj, subset = assigned_celltype %in% c("Oligodendrocytes_Mature", "Oligodendrocyte_Precursors"))
```

```{r}
oligo_obj@meta.data$Tumor_Periphery <- ifelse(
  oligo_obj@meta.data$Location == "Tumor",
  "Tumor",
  ifelse(
    oligo_obj@meta.data$Location == "Periphery",
    "Periphery",
    ifelse(
      oligo_obj@meta.data$Location == "Distant",
      "Distant",
      "Other"
    )
  )
)
```

```{r warning=FALSE}
oligo_obj <- SCTransform(oligo_obj, verbose = FALSE)
oligo_obj <- PrepSCTFindMarkers(oligo_obj)
oligo_obj <- RunPCA(oligo_obj, dims = 1:30)
# Add Dimensionality Reduction (UMAP)
oligo_obj <- RunUMAP(oligo_obj, dims = 1:30)
```

```{r}
# Compare Tumor vs. Periphery
Idents(oligo_obj) <- "Tumor_Periphery"
tumor_vs_periphery_oligo_genes <- FindMarkers(oligo_obj, ident.1 = "Tumor", ident.2 = "Periphery", min.pct = 0.25, logfc.threshold = 0.25)
```

```{r}
# Compare Tumor vs. Distant
tumor_vs_distant_oligo_genes <- FindMarkers(oligo_obj, ident.1 = "Tumor", ident.2 = "Distant", min.pct = 0.25, logfc.threshold = 0.25)
```

```{r}
# Compare Periphery vs. Distant
periphery_vs_distant_oligo_genes <- FindMarkers(oligo_obj, ident.1 = "Periphery", ident.2 = "Distant", min.pct = 0.25, logfc.threshold = 0.25)
```

```{r}
# Visualization and Interpretation (Oligodendrocytes)

# Function to visualize and interpret DE results for oligo_obj
visualize_de_results_oligo <- function(de_genes, comparison_name, oligo_obj, top_n_heatmap = 20, top_n_features = 3) {
  print(paste("Results for:", comparison_name))

  # View top differentially expressed genes
  head(de_genes)

  # Filter for significant and substantial changes
  significant_genes <- de_genes %>%
    filter(p_val_adj < 0.05, abs(avg_log2FC) > 0.5) %>%
    arrange(desc(avg_log2FC))

  print(significant_genes)

  # Check if significant_genes is empty
  if (nrow(significant_genes) == 0) {
    print("No significant genes found with current filtering criteria.")
    return(NULL) # Exit the function if no genes are found
  }

  # Volcano Plot (Oligodendrocytes)
  volcano_plot_oligo <- EnhancedVolcano::EnhancedVolcano(de_genes,
    lab = rownames(de_genes),
    x = "avg_log2FC",
    y = "p_val_adj",
    pCutoff = 0.05,
    FCcutoff = 0.5,
    title = paste("Volcano Plot:", comparison_name),
    subtitle = "Значущі гени виділено",
    caption = paste0("Всього змінних = ", nrow(de_genes)),
    col = c("grey30", "forestgreen", "royalblue", "firebrick2"),
    colAlpha = 0.8,
    boxedLabels = TRUE,
    drawConnectors = TRUE,
    widthConnectors = 0.5,
    colConnectors = "grey50",
    max.overlaps = 20,
    legendPosition = "topright",
    legendLabSize = 10,
    legendIconSize = 4.0,
    titleLabSize = 14,
    subtitleLabSize = 12,
    captionLabSize = 10,
    xlab = bquote(~Log[2]~ " зміна експресії"),
    ylab = bquote(~-Log[10]~italic(P)),
    pointSize = 1.5
  )
  print(volcano_plot_oligo)

  # Feature Plots (Top N genes)
  if (nrow(significant_genes) > 0) {
    tryCatch({
      reductions <- Reductions(oligo_obj)
      if ("umap_harmony" %in% reductions) {
        reduction_to_use <- "umap_harmony"
      } else if (length(reductions) > 0) {
        reduction_to_use <- reductions[1]
      } else {
        print("No reductions found in oligo_obj.")
        return(NULL)
      }
      top_feature_genes <- head(rownames(significant_genes), top_n_features)
      feature_plot_oligo <- FeaturePlot(oligo_obj, features = top_feature_genes, reduction = reduction_to_use, cols = c("lightgrey", "red"), pt.size = 0.5)
      print(feature_plot_oligo)
    }, error = function(e) {
      print(paste("Помилка при створенні FeaturePlot:", e$message))
    })
  }


# Heatmap Visualization (з групуванням за поточними Idents) для oligo_obj
if (nrow(significant_genes) > 0) {
  # Вибираємо топ N значущих генів
  top_genes <- head(rownames(significant_genes), top_n_heatmap)

  # Масштабуємо дані за допомогою ScaleData для oligo_obj
  oligo_obj <- ScaleData(oligo_obj, features = top_genes, verbose = FALSE)

  # Створюємо хітмапу (групування за поточними Idents) для oligo_obj
  heatmap_plot <- DoHeatmap(object = oligo_obj, features = top_genes, assay = "SCT", slot = "scale.data") +
    ggplot2::ggtitle(paste("Heatmap (Scaled):", comparison_name)) +
    ggplot2::theme(plot.title = ggplot2::element_text(color = "white", size = 2, hjust = 0, vjust = 0)) 
  print(heatmap_plot)
}

  gene_list <- rownames(significant_genes)
  if (length(gene_list) > 0) {
    entrez_ids <- bitr(gene_list, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Hs.eg.db)

    if (nrow(entrez_ids) > 0) {
      # GO Enrichment
      go_results <- enrichGO(gene = entrez_ids$ENTREZID,
        OrgDb = org.Hs.eg.db,
        ont = "BP",
        pAdjustMethod = "BH",
        pvalueCutoff = 0.01,
        qvalueCutoff = 0.05
      )

if (!is.null(go_results)) {
  tryCatch({
    print(enrichplot::dotplot(go_results, showCategory = 10, title = paste("GO Enrichment:", comparison_name)) +
            ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8, angle = 0, hjust = 1, vjust = 0.5), 
                           plot.title = ggplot2::element_text(hjust = 0.5))) 
  }, error = function(e) {
    print(paste("Недостатньо даних для створення dotplot або структура go_results не відповідає очікуваній.", e$message))
  })
}

# Hallmark Gene Sets
library(msigdbr)
hallmark_gene_sets <- msigdbr(species = "Homo sapiens", category = "H")

gene_ranks_symbol <- de_genes$avg_log2FC
names(gene_ranks_symbol) <- rownames(de_genes)
entrez_map <- bitr(names(gene_ranks_symbol), fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Hs.eg.db)

# Створюємо ранжований список з Entrez ID як іменами
gene_ranks_entrez_named <- gene_ranks_symbol[entrez_map$SYMBOL]
names(gene_ranks_entrez_named) <- entrez_map$ENTREZID

# Видаляємо записи з NA в іменах (Entrez ID)
gene_ranks_entrez_filtered <- gene_ranks_entrez_named[!is.na(names(gene_ranks_entrez_named))]

# Сортуємо ранжований список
gene_ranks_entrez_filtered <- sort(gene_ranks_entrez_filtered, decreasing = TRUE)

hallmark_results <- GSEA(gene_ranks_entrez_filtered,
                         TERM2GENE = hallmark_gene_sets %>% dplyr::select(gs_name, entrez_gene),
                         pvalueCutoff = 0.05
)
if (!is.null(hallmark_results) && nrow(hallmark_results) > 0) {
  tryCatch({
    print(enrichplot::dotplot(hallmark_results, showCategory = 10, title = paste("Hallmark Enrichment:", comparison_name)) +
            ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8, angle = 0, hjust = 1, vjust = 0.5), 
                           plot.title = ggplot2::element_text(hjust = 0.5))) 
  }, error = function(e) {
    print(paste("Помилка при створенні dotplot (Hallmark):", e$message))
  })
} else {
  print("Немає значущих результатів збагачення Hallmark gene sets для відображення.")
}
    }
  }
}
```

```{r warning=FALSE}
# Apply visualization function to each comparison (Oligodendrocytes) with custom top gene numbers
visualize_de_results_oligo(tumor_vs_periphery_oligo_genes, "Tumor vs. Periphery (Oligodendrocytes)", oligo_obj, top_n_heatmap = 15, top_n_features = 4)
visualize_de_results_oligo(tumor_vs_distant_oligo_genes, "Tumor vs. Distant (Oligodendrocytes)", oligo_obj, top_n_heatmap = 15, top_n_features = 4)
visualize_de_results_oligo(periphery_vs_distant_oligo_genes, "Periphery vs. Distant (Oligodendrocytes)", oligo_obj, top_n_heatmap = 15, top_n_features = 4)
```
